## mTOR in Penile Cancer
Last update: ```r {date()}```

```{r global_options}
library(knitr)
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "asis")
```

```{r DataSet}
# Loading the working dataset
Data <- read.csv("PenisDataset.csv")
attach(Data)

# Loading scripts and libraries
source("RFUN/NicePlots.R")
source("RFUN/NiceTables.R")
```

### Descriptive Analysis
In this section we describe the clinicopathologic and outcome features of the dataset. Numerical variables were described using histograms and boxplots. Categorical variables were described using bar plots.

***

#### Patient's age
```{r DescriptiveAge}
var <- age
t <- "Patient's age, years"
plot.numerical(var, label = t)
descriptive.numerical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Patient's race
```{r DescriptiveRace}
var <- race
t <- "Patient's race"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Histologic subtype
```{r DescriptiveSubtype}
var <- subtype
t <- "Histologic Subtype"
plot.categorical(var, align = "h", left = 5)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Anatomical site
```{r DescriptiveSite}
var <- site
t <- "Anatomical Site"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Histologic grade
```{r DescriptiveGrade}
var <- grade
t <- "Histologic Grade"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Tumor thickness
```{r DescriptiveThick}
thick <- factor(thick, levels = c("Less than 5 mm", "5-10 mm", "More than 10 mm"))
var <- thick
t <- "Tumor Thickness"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Lymphovascular invasion
```{r DescriptiveLVI}
var <- lvi
t <- "Lymphovascular Invasion"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Perineural invasion
```{r DescriptivePNI}
var <- pni
t <- "Perineural Invasion"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Anatomical level
```{r DescriptiveLevel}
var <- level
t <- "Anatomical Level"
plot.categorical(var, align = "h", left = 7)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Invasion of penile urethra
```{r DescriptiveUrethra}
var <- urethra
t <- "Invasion of Penile Urethra"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Inguinal lymph node metastasis
```{r DescriptiveNodes}
var <- nodes
t <- "Inguinal LN metastasis"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Date of lymph node dissection
```{r DescriptiveDateLN}
var <- date.lymph.mo
t <- "Months to LN dissection"
plot.numerical(var, label = t)
descriptive.numerical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

#### Type of inguinal dissection
```{r DescriptiveDissection}
var <- dissection.type
t <- "Type of Inguinal Dissection"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### pT Stage
```{r DescriptiveT}
tnm.tumor <- factor(tnm.tumor, levels = c("Ta", "T1a", "T1b", "T2", "T3", "T4"))
var <- tnm.tumor
t <- "pT Stage"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### pN Stage
```{r DescriptiveN}
var <- tnm.nodes
t <- "pN Stage"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Clinical stage
```{r DescriptiveStage}
var <- stage
t <- "Clinical Stage"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Primary treatment
```{r DescriptiveTreatment}
var <- treatment
t <- "Primary Treatment"
plot.categorical(var, align = "h", left = 10)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Adjuvant chemotherapy
```{r DescriptiveChemo}
var <- chemo
t <- "Adjuvant Chemotherapy"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Adjuvant radiotherapy
```{r DescriptiveRadio}
var <- radio
t <- "Adjuvant Radiotherapy"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Tumor recurrence
```{r DescriptiveRecurrence}
var <- recurrence
t <- "Tumor Recurrence"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Time to tumor recurrence
```{r DescriptiveRecurrenceTime}
var <- date.recurrence.mo
t <- "Months to Tumor Recurrence"
plot.numerical(var, label = t)
descriptive.numerical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Tumor progression
```{r DescriptiveProgression}
var <- progression
t <- "Tumor Progression"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Time to tumor recurrence
```{r DescriptiveProgressionTime}
var <- date.progression.mo
t <- "Months to Tumor Progression"
plot.numerical(var, label = t)
descriptive.numerical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Follow-Up
```{r DescriptiveFU}
var <- fu.last.mo
t <- "Follow-Up (Months)"
plot.numerical(var, label = t)
descriptive.numerical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Outcome
```{r DescriptiveOutcome}
var <- outcome
t <- "Patient's Outcome"
plot.categorical(var, align = "h", left = 9)  
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Overall mortality
```{r DescriptiveDead}
var <- dead
t <- "Overall Mortality"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***