## mTOR in Penile Cancer: PTEN expression
Last update: ```r {date()}```

```{r global_options}
library(knitr)
opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, results = "asis")
```

```{r PTEN_DataSet}
# Loading the working dataset
Data <- read.csv("PenisDataset.csv")
attach(Data)

# Loading scripts and libraries
source("RFUN/NicePlots.R")
source("RFUN/NiceTables.R")
```

### Description
In this section we evaluate the usefulness of `PTEN expression` for predicting outcome, including tumor recurrence, tumor progression, distant metastasis, overall mortality, and cancer-related mortality. The analyses include association tests (Fisher's exact test), survival analysis (Kaplan-Meier method and log-rank test), unconditional binary logistic regression (for estimating odds ratios), and Cox's proportional hazards regression (for estimating hazard ratios). 

__Methodology:__ Immunohistochemical expression was categorized in 2 levels (decreased/loss and maintained) using the `lower tertile` as the cutoff point.

```{r PTEN_Levels}
marker.cut <- quantile(PTEN, probs = 0.33, na.rm = TRUE)
marker <- ifelse(PTEN < marker.cut, c("Decreased/Loss PTEN"), c("Maintained PTEN"))
marker <- factor(marker, levels = c("Maintained PTEN", "Decreased/Loss PTEN"))
```
***

#### Immunohistochemical expression by H-scores

```{r PTEN_DescriptiveH}
var <- PTEN
t <- "H-score"
plot.numerical(var, label = t)
descriptive.numerical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

***

#### Immunohistochemical expression by levels

```{r PTEN_DescriptiveLevels}
var <- marker
t <- "Expression Levels"
plot.categorical(var, align = "v", xlab = t)
descriptive.categorical(var)
```

_Number of missing cases:_ `r sum(is.na(var))` cases.

### Association of Immunohistochemical Expression with Outcome
In this subsection we evaluate the association of levels of immunohistochemical expression with outcome (including tumor recurrence, tumor progression, distant metastasis, overall mortality, and cancer-related mortality) using the Fisher's exact test.

***

#### Tumor recurrence
```{r PTEN_AssociationRecurrence}
outcome <- recurrence
t <- "Tumor Recurrence"
plot.categorical.group(marker, outcome, align = "v", xlab = t)
descriptive.categorical.group(marker, outcome)

```

***

#### Tumor progression
```{r PTEN_AssociationProgression}
outcome <- progression
t <- "Tumor Progression"
plot.categorical.group(marker, outcome, align = "v", xlab = t)
descriptive.categorical.group(marker, outcome)

```

***

#### Distant metastasis
```{r PTEN_AssociationDistantMetastasis}
outcome <- mets
t <- "Distant Metastasis"
plot.categorical.group(marker, outcome, align = "v", xlab = t)
descriptive.categorical.group(marker, outcome)

```

***

#### Overall mortality
```{r PTEN_AssociationOverallMortality}
outcome <- dead
t <- "Overall Mortality"
plot.categorical.group(marker, outcome, align = "v", xlab = t)
descriptive.categorical.group(marker, outcome)

```

***

#### Cancer-related mortality
```{r PTEN_AssociationCancerMortality}
outcome <- died.disease
t <- "Cancer-Related Mortality"
plot.categorical.group(marker, outcome, align = "v", xlab = t)
descriptive.categorical.group(marker, outcome)

```

***

### Survival Analysis by Immunohistochemical Expression
In this subsection we plot survival curves for levels of immunohistochemical expression by outcome (including tumor recurrence, tumor progression, distant metastasis, overall mortality, and cancer-related mortality) using the Kaplan-Meier method. Survival curves were compared using the log-rank (Mantel-Cox) test.

***

```{r PTEN_SurvivalRecurrence, fig.width = 10, fig.height = 7}
# Tumor recurrence
outcome <- recurrence
fu <- date.recurrence.mo
t <- "Tumor Recurrence"
survival.plot(marker, fu, outcome, title = t, ylim = c(-0.05, 1))

```

***

```{r PTEN_SurvivalProgression, fig.width = 10, fig.height = 7}
# Tumor progression
outcome <- progression
fu <- date.progression.mo
t <- "Tumor Progression"
survival.plot(marker, fu, outcome, title = t)

```

***

```{r PTEN_SurvivalDistantMetastasis, fig.width = 10, fig.height = 7}
# Distant metastasis
outcome <- mets
fu <- date.met.mo
t <- "Distant Metastasis"
survival.plot(marker, fu, outcome, title = t, ylim = c(0, 1.099))

```

***

```{r PTEN_SurvivalOverallMortality, fig.width = 10, fig.height = 7}
# Overall mortality
outcome <- dead
fu <- fu.last.mo
t <- "Overall Mortality"
survival.plot(marker, fu, outcome, title = t)

```

***

```{r PTEN_SurvivalCancerMortality, fig.width = 10, fig.height = 7}
# Cancer-related mortality
outcome <- died.disease
fu <- fu.last.mo
t <- "Cancer-Related Mortality"
survival.plot(marker, fu, outcome, title = t, ylim = c(0.4, 1))

```

***

### Unconditional Binary Logistic Regression
In this subsection we evaluate the usefulness of clinicopathologic features and immunohistochemical expression in predicting outcome (including tumor recurrence, tumor progression, distant metastasis, overall mortality, and cancer-related mortality) using unconditional binary logistic regression models. For each model we report the odds ratio (OR) with its 95% confidence intervals (lower and upper limits), raw P values, and P values adjusted using Hommel's correction.

```{r PTEN_VariablesOR}
# Recoding new variables for regression models
older <- ifelse(age > median(age), c("Older"), c("Younger"))
caucasian <- ifelse(race == "Caucasian", c("Yes"), c("No"))
glans <- ifelse(site == "Glans", c("Yes"), c("No"))
lp <- ifelse(site == "Lamina Propria", c("Yes"), c("No"))
highgrade <- ifelse(grade == "Grade 3", c("Yes"), c("No"))
deep <- ifelse(thick == "More than 10 mm", c("Yes"), c("No"))
highpt <- ifelse(as.numeric(tnm.tumor) > 2, c("Yes"), c("No"))
highpn <- ifelse(as.numeric(tnm.nodes) > 1, c("Yes"), c("No"))
highstage <- ifelse(as.numeric(stage) > 2, c("Yes"), c("No"))
# Creating a list of predictors for regression models
predictors <- list("Patient's age > median" = older,
                   "Caucasian race" = caucasian,
                   "Glans location" = glans,
                   "Invasion limited to lamina propria" = site,
                   "High grade tumor (grade 3)" = highgrade,
                   "Deep invasive tumor (> 10 mm)" = deep,
                   "Invasion of penile urethra" = urethra,
                   "Lymphovascular invasion" = lvi,
                   "Perineural invasion" = pni,
                   "HPV infection" = hpv.pos,
                   "High pT stage (> pT1)" = highpt,
                   "High pN stage (> pN0)" = highpn,
                   "High clinical stage (> Stage II)" = highstage,
                   "Prophylactic inguinal lymph node dissection" = dissection.type,
                   "Metastatic inguinal lymph nodes" = nodes,
                   "PTEN expression decreased/lost" = marker
                   )
# Creating the variable labels for regression tables
varlabels <- names(predictors)
# Creating the variable of follow-up time for regression tables
fu <- fu.last.mo
```

***

#### Tumor recurrence
```{r PTEN_RecurrenceOR}
outcome <- recurrence
logistic.table(outcome, predictors, varlabels)
```

***

#### Tumor progression
```{r PTEN_ProgressionOR}
outcome <- progression
logistic.table(outcome, predictors, varlabels)
```

***

#### Distant metastasis
```{r PTEN_DistantMetastasisOR}
outcome <- mets
logistic.table(outcome, predictors, varlabels)
```

***

#### Overall mortality
```{r PTEN_OverallMortalityOR}
outcome <- dead
logistic.table(outcome, predictors, varlabels)
```

***

#### Cancer-related mortality
```{r PTEN_CancerMortalityOR}
outcome <- died.disease
logistic.table(outcome, predictors, varlabels)
```

***

### Cox's Proportional Hazards Regression
In this subsection we evaluate the usefulness of clinicopathologic features and immunohistochemical expression in predicting outcome (including tumor recurrence, tumor progression, distant metastasis, overall mortality, and cancer-related mortality) using unadjusted Cox's proportional hazard regression models. For each model we report the hazard ratio (HR) with its 95% confidence intervals (lower and upper limits), raw P values, and P values adjusted using Hommel's correction.

***

#### Tumor recurrence
```{r PTEN_RecurrenceHR}
outcome <- recurrence
cox.table(outcome, fu, predictors, varlabels)
```

***

#### Tumor progression
```{r PTEN_ProgressionHR}
outcome <- progression
cox.table(outcome, fu, predictors, varlabels)
```

***

#### Distant metastasis
```{r PTEN_DistantMetastasisHR}
outcome <- mets
cox.table(outcome, fu, predictors, varlabels)
```

***

#### Overall mortality
```{r PTEN_OverallMortalityHR}
outcome <- dead
cox.table(outcome, fu, predictors, varlabels)
```

***

#### Cancer-related mortality
```{r PTEN_CancerMortalityHR}
outcome <- died.disease
cox.table(outcome, fu, predictors, varlabels)
```

***
